Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-05-17 manuscript/manuscript.tex
Modifications in conclusions

We have developed a new methodology that computes the gravitational generated 
by a tesseroid with variable density in depth.
It numerically solves the integrals that define the gravitational potential, 
its gradient and the Marussi tensor through the Gauss-Legendre Quadrature 
(GLQ), which essentially consists in approximating the gravitational fields of 
the tesseroids with the effect of point masses located on the scaled nodes of 
the Legendre polynomials.

@@

The former one divides the tesseroid by half if the ratio of the distance to 
the computation point and its size is lower than a predefined distance-size 
ratio $D$.
It's intended to minimise the error when the computation point is close to the 
tesseroid.
We can summarise that the distance-size ratio $D$ controls the accuracy of the 
method: a higher value of $D$ generates more divisions of the tesseroids, thus 
more point masses, what leads to a more precise approximation.
Nevertheless, this algorithm is not sufficient to guarantee the accuracy of 
the method in case of tesseroid with variable density.

For this reason we have developed a density-based discretization algorithm that 
divides the tesseroid on the radius at which the \emph{maximum variation} of 
the density function takes place, only if a certain inequation that relates 
this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
Therefore the number of subdivisions is indirectly controlled by this delta 
ratio $\delta$, and thus the accuracy of the computation.

Because there is no direct relation between the values assigned to $D$ and 
$\delta$ with the error of the computation, we had to empirically determine 
the minimum value of $D$ and the maximum value of $\delta$ that produce an 
acceptable accuracy for each gravitational field, while maintaining an 
efficient computation.
In order to determine default values of $D$ and $\delta$ we have chosen two 
density functions: a linear and an exponential one.

@@

The exponential density case is more complex, due to the fact that both $D$ 
and $\delta$ must be determined simultaneously.
We have explored point in the $D$, $\delta$ space and computed the percentage 
difference between the analytical solutions for a spherical shell and the 
numerical model.
We had to take into account that the $D$ values cannot be lower than the ones 
determined for the linear density case due to the fact that they must 
guarantee acceptable accuracy for every density function, including the linear 
one.
From these comparisons we have concluded that using the same $D$ values 
obtained in the linear density scenario, a $\delta = 0.2$ is needed for the 
computation of the potential and its gradient components, while a $\delta = 
0.01$ must be used for the Marussi tensor components in order to ensure a 
0.1\% precision.

@@

Finally, we have shown a real data example: we have computed the gravity 
fields generated by the Neuqu\'en sedimentary basin through a tesseroid model 
with exponential density with depth on a regular grid at 50km height.

>>>>

We have developed a new methodology that computes the gravitational fields 
generated by a tesseroid with continuously variable density in depth.
It numerically solves the integrals that define the gravitational potential, 
its gradient and the Marussi tensor components through the Gauss-Legendre 
Quadrature (GLQ), which essentially consists in approximating the 
gravitational fields of the tesseroids with the effect of point masses located 
on nodes of the Quadrature.

@@

The former one divides the tesseroid by half if the ratio of the distance to 
the computation point and its size is lower than a predefined distance-size 
ratio $D$.
This algorithm is introduced in order to minimise the error when the 
computation point is close to the tesseroid, in which case the effect of the 
point masses becomes more noticeable and the approximation lose precision.
Following this idea, we can summarise that the distance-size ratio $D$ 
controls the accuracy of the method: a higher value of $D$ generates more 
divisions of the tesseroids, thus more point masses, what leads to a more 
precise approximation.
Nevertheless, this algorithm is not sufficient to guarantee the accuracy of 
the method in case of tesseroid with variable density.

For this reason we have developed a density-based discretization algorithm that 
divides the tesseroid on the radius at which the \emph{maximum variation} of 
the density function takes place, only if a certain inequation that relates 
this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
The the number of subdivisions is indirectly controlled by this delta 
ratio $\delta$, and thus the accuracy of the computation.
This new algorithm is intended to minimise the error due to the inability of 
the GLQ to produce precise approximations in case of highly varying density: 
the information about the density function is summed up with the values that 
it assumes at the node of the quadrature.
Increasing the subdivisions of the tesseroid where the \emph{maximum 
variation} of the density takes place becomes an efficient way to reduce this 
kind of error.

Because there is no direct relation between the values assigned to $D$ and 
$\delta$ with the error of the computation, we had to empirically determine 
the minimum value of $D$ and the maximum value of $\delta$ that produce an 
acceptable accuracy for each gravitational field, while maintaining an 
efficient computation.
The default values of $D$ and $\delta$ must produce accurate computations for 
most continuous and smooth densities, that's why we have chosen two density 
functions in order to establish them: a linear and an exponential one.

@@

The exponential density case is more complex, due to the fact that both $D$ 
and $\delta$ must be determined simultaneously.
We have explored regular gridded points in the $D$, $\delta$ space and 
computed the percentage difference between the analytical solutions for a 
spherical shell and the numerical model.
We had to take into account that the $D$ values cannot be lower than the ones 
determined for the linear density case due to the fact that they must 
guarantee acceptable accuracy for every continuous density function, including 
the linear one.
From these comparisons we have concluded that the more efficient computations 
take places when the $D$ values obtained in the linear density scenario are 
used, i.e. the minimum ones.
This can be understood in terms of which kind of discretization produces more 
subdivisions: a higher $D$ may produce more discretizations than a lower 
$\delta$ because the first one divides the tesseroid in the radial, longitude 
and latitude directions, while the latter only divides it in the radial 
direction.
In conclusion, we found out that keeping the $D$ values as the ones obtained 
for the linear density case, a $\delta = 0.2$ is needed for the computation of 
the potential and its gradient components, while a $\delta = 0.01$ must be 
used for the Marussi tensor components in order to ensure a 0.1\% precision.

@@

Finally, we have shown how this new method performs under a real data example: 
we have computed the gravitational fields generated by the Neuqu\'en 
sedimentary basin through a tesseroid model with an increasing exponential 
density in depth on a regular grid at 50km height.
